\documentclass[11pt]{article}

% Edit this customize for an instructor

\newcommand{\instructorpronoun}[1]{his}

% Use this when displaying a new command

\newcommand{\command}[1]{``\lstinline{#1}''}
\newcommand{\program}[1]{\lstinline{#1}}
\newcommand{\url}[1]{\lstinline{#1}}
\newcommand{\channel}[1]{\lstinline{#1}}
\newcommand{\option}[1]{``{#1}''}
\newcommand{\step}[1]{``{#1}''}

\usepackage{pifont}
\newcommand{\checkmark}{\ding{51}}
\newcommand{\naughtmark}{\ding{55}}

\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  columns=flexible,
  breaklines=true
}

% Define the headers and footers

\usepackage{fancyhdr}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}

\pagestyle{fancy}

\fancyhf{}
\rhead{Computer Science 280}
\lhead{Syllabus}
\rfoot{Page \thepage}
\lfoot{Fall 2017}

% Use elastic spacing around the headers

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\syllabustitle}[1]
{
  \begin{center}
    \begin{center}
      \bf
      CMPSC 280\\Software Engineering\\
      Fall 2017\\
      \medskip
    \end{center}
    \bf
    #1
  \end{center}
}

\begin{document}

\thispagestyle{empty}

\syllabustitle{Syllabus}

\subsection*{Course Instructor}
Dr.\ Gregory M.\ Kapfhammer\\
\noindent Office Location: Alden Hall 108 \\
\noindent Office Phone: +1 814-332-2880 \\
\noindent Email: \url{gkapfham@allegheny.edu} \\
\noindent Twitter: \url{@GregKapfhammer} \\
\noindent Web Site: \url{http://www.cs.allegheny.edu/sites/gkapfham/}

\subsection*{Instructor's Office Hours}

\begin{itemize}
  \itemsep0em

  \item Monday: 2:30 pm--3:30 pm (15 minute time slots)

  \item Wednesday: 10:00 am--11:00 am and 2:30 pm--3:30 pm (15 minute time slots)

  \item Thursday: 11:00 pm--12:00 noon (15 minute time slots)

  \item Friday: 10:00 am--11:00 am and 2:30 pm--3:30 pm (15 minute time slots)

\end{itemize}

\noindent To schedule a meeting with me during my office hours, please visit my web site and click the ``Schedule'' link
in the top right-hand corner. Now, you can browse my office hours or schedule an appointment by clicking the correct
link and then reserving an open time slot. Students are also encouraged to post appropriate questions to a channel in
Slack, which is available at \url{https://CMPSC280Fall2017.slack.com/}, and monitored by the instructor and the
teaching assistants.

\subsection*{Course Meeting Schedule}

Lecture, Discussion, and Group Work: Monday, Wednesday, Friday, 1:30 pm--2:20 pm \\
Laboratory Session: Tuesday, 2:30 pm--4:20 pm \\
Final Examination: Friday, December 11, 2015 at 9:00 am

\subsection*{Course Description}

\begin{quote}

  A human-centric study of the principles used during the engineering of high-quality software systems. In addition to
  examining the human behaviors and social processes undergirding software development methodologies, students participate
  in teams tasked with designing, developing, and delivering a significant software application for a customer. During a
  weekly laboratory session, students use state-of-the-art software engineering, management, and communication tools to
  complete projects, reporting on their results through both written reports and oral presentations. Prerequisite: CMPSC
  112. Distribution Requirements: SB, SP.

\end{quote}

\subsection*{Course Objectives}

The process of developing software involves the application of a number of interesting theories, tools, techniques, and
methodologies. In this class we will explore the phases of the major software development life cycles and examine the
tools, concepts, challenges, and open questions associated with each phase. Throughout the semester, we will examine
the interplay between the theory and practice of software development. Also, we will delve into the details of software
specification, design, implementation, testing, and maintenance through a discussion of book chapters and articles from
the software engineering and software testing literature. Along with learning more about how to effectively work in a
team of diverse software developers, students will enhance their ability to write and present ideas about software in a
clear, concise, and compelling fashion. Students will develop an understanding of the fascinating connections
between computer science and software engineering and other disciplines in the social and natural sciences and the
humanities. Students also will gain practical software development experience in laboratory sessions and a final
project.

\subsection*{Performance Objectives}

At the completion of this class, a student should be aware of the fundamental challenges associated with software
development. Furthermore, students should be comfortable with a wide array of concepts, methodologies, techniques, and
tools that they can apply to the problem of developing large software systems. However, a successful student will
emerge with more than an understanding of the tools (e.g., text editors, compilers, debuggers, integrated development
environments, and version control systems) that a software engineer uses. A student also should have a fundamental
understanding of the major software life cycles and the activities that take place in each phase of these life cycles.
Finally, a student should have a basic understanding of some of the current research and the open questions in the field
of software engineering. After completing this class, a student should be equipped for further graduate study in the
fields of computer science and software engineering. The student should also be able to participate in real-world
software development projects by adeptly using cutting-edge software tools and working with a team of diverse
developers.

\subsection*{Required Textbooks}

% Shari Lawrence Pfleeger and Joanne M. Atlee
%   Software Engineering: Theory and Practice (Fourth Edition)
%   ISBN-10: 0136061699
%   ISBN-13: 978-0136061694
%   Prentice Hall
%   Status: Required
%   25 copies

\noindent{\em Software Engineering: Theory and Practice}. Shari Lawrence Pfleeger and Joanne M. Atlee.
Fourth Edition, ISBN-10: 0136061699, ISBN-13: 978--0136061694, 792 pages, 2010. \\
(References to the textbook are abbreviated as ``SETP'' in the syllabus and on the Web site).

% The Mythical Man-Month: Essays on Software Engineering, Anniversary
%  Edition (2nd Edition) [Paperback]
%  Author: Frederick P. Brooks
%  Publisher: Addison-Wesley Professional; Anniversary edition (August 12, 1995)
%  ISBN-10: 0201835959
%  ISBN-13: 978-0201835953
%  Status: Required
%  25 copies

\noindent{\em The Mythical Man Month}. Frederick P.\ Brooks, Jr.
Second Edition, ISBN-10: 0201835959, ISBN-13: 978--0201835953, 336 pages, 1995. \\
(References to the textbook are abbreviated as ``MMM'' in the syllabus and on the Web site).

\noindent
Students who want to improve their technical writing skills may consult the following books.

\noindent{\em BUGS in Writing: A Guide to Debugging Your Prose}. Lyn Dupr\'e. Second Edition,  ISBN-10: 020137921X,
ISBN-13: 978--0201379211, 704 pages, 1998.

\noindent{\em Writing for Computer Science}. Justin Zobel. Second Edition, ISBN-10: 1852338024, ISBN-13:
978--1852338022, 270 pages, 2004.

\noindent
Along with reading the required books, you may be asked to study many additional articles from a wide variety of
conference proceedings, journals, and the popular press.

\subsection*{Class Policies}

\subsubsection*{Grading}

The grade that a student receives in this class will be based on the following categories. All percentages are
approximate and, if the need to do so presents itself, it is possible for the course instructor to change the assigned
percentages during the academic semester.

\begin{center}
  \begin{tabular}{ll}
    Class Participation        & 5\%  \\
    SEED Project Participation & 5\%  \\
    First Examination          & 15\% \\
    Second Examination         & 15\% \\
    Final Examination          & 20\% \\
    Laboratory Assignments     & 25\% \\
    Final Project              & 15\%
  \end{tabular}
\end{center}

\vspace*{-.1in}
\noindent
These grading categories have the following definitions:
\vspace*{-.1in}

\begin{itemize}

  \item {\em Class Participation\/}: All students are required to actively participate during all of the course
    sessions. Your participation will take forms such as answering questions about the reading assignments, asking
    constructive questions of group members, giving presentations, and leading a discussion. You also must regularly
    participate in the discussions in the course's Slack team. A student may request feedback on and will receive a
    final grade for this category.

  \item {\em SEED Project Participation\/}: As part of an ongoing project that publishes educational discussions with
    software engineers, every week I will publish, on my blog available at
    \url{http://www.cs.allegheny.edu/sites/gkapfham/blog/}, an interview with a software developer in industry. At the
    start of the practical session on Friday of each week, every student should read the blog post and tweet a response
    to that week's interview. Along with linking to the web site of the blog post and using the hashtag \url{#SEED},
    your tweet should share what you learned from the interviewee. Interested students should also connect with the
    interviewee through LinkedIn.

  \item {\em First and Second Examinations\/}: The first and second interim examinations will cover all of the material
    in their associated module(s). While the second examination is not cumulative, it will assume that a student has a
    basic understanding of the material that was the focus of the first examination. The date for the first and second
    examinations will be announced at least one week in advance of the scheduled date. Unless prior arrangements are
    made with the course instructor, all students will be expected to take these examinations on the scheduled date and
    complete the tests in the stated period of time.

  \item {\em Final Examination\/}: The final examination is a three-hour cumulative test. By enrolling in this course,
    students agree that, unless there are extenuating circumstances, they will take the final examination at the time
    stated on the first page of the syllabus.

  \item {\em Laboratory and Homework Assignments\/}: These assignments invite students to explore the concepts, tools,
    and techniques that are associated with different phases of the software development life cycle. All of the
    laboratory assignments require the use of the provided tools to design, implement, test, maintain, and document
    programs that solve important problems. To ensure that students are ready to develop software in both other classes
    at Allegheny College and after graduation, the instructor will assign individuals to teams for each of the
    laboratory assignments. Unless specified otherwise, each laboratory assignment will be due at the beginning of the
    next laboratory session. Many of the laboratory assignments in this course will expect students to give both a
    presentation and a demonstration of the software that they specified, designed, implemented, tested, and documented.

  \item {\em Final Project\/}: This project will present you with the description of a problem and ask you to implement
    a full-featured solution using one or more programming languages and a wide variety of software development tools.
    The final project in this class will require you to apply all of the knowledge and skills that you have accumulated
    during the semester to solve a problem and, whenever possible, make your solution publicly available as a free and
    open-source tool. The project will invite you to draw upon both your communication and problem solving skills and
    your knowledge of programming languages and software engineering tools. The final project will be completed in
    groups assigned by the course instructor.

\end{itemize}

\subsubsection*{Assignment Submission}

All assignments will have a stated due date. The printed version of the assignment is to be turned in at the beginning
of the class on that due date; the printed materials must be dated and signed with the Honor Code pledge of all the
students in the group.  When the printed version is submitted, the electronic version of the assignment also must be
made available to the course instructor in a version control repository. Late assignments will be accepted for up to one
week past the assigned due date with a 15\% penalty. All late assignments must be submitted at the beginning of the
session that is scheduled one week after the due date. Unless special arrangements are made with the course instructor,
no assignments will be accepted after the late deadline. In addition to submitting the required deliverables for any
assignment completed in a group, students must turn in a one-page document that describes each group member's
contribution to the submitted deliverables.

\subsubsection*{Attendance}

It is mandatory for all students to attend the class and laboratory sessions and all group project meetings. If, due to
extenuating circumstances, you will not be able to attend one of these events, then, whenever possible, please see the
instructor at least one week in advance to describe your situation. Students who miss more than five unexcused classes,
laboratory sessions, or group project meetings will have their final grade in the course reduced by one letter grade.
Students who miss more than ten of the aforementioned events will automatically fail the course.

\subsubsection*{Use of Laboratory Facilities}

Throughout the semester, we will experiment with many different tools that software engineers use during the phases of
the software development life cycle.  The course instructor and the department's systems administrator have invested a
considerable amount of time to ensure that our laboratories support the completion of both the laboratory assignments and the
final project.  To this end, students are required to complete all assignments and the final project while using the
department's laboratory facilities. The course instructor and the systems administrator will only be able to devote a
very limited amount of time to the configuration of a student's personal computer.

\subsubsection*{Class Preparation}

In order to minimize confusion and maximize learning, students must invest time to prepare for class discussions and
lectures.  During the class periods, the course instructor will often pose demanding questions that could require group
discussion, the creation of a program or test suite, a vote on a thought-provoking issue, or a group presentation.
Only students who have prepared for class by reading the assigned material and reviewing the current assignments will be
able to effectively participate in these discussions.  More importantly, only prepared students will be able to acquire
the knowledge and skills that are needed to be successful in both this course and the field of software development.  In
order to help students remain organized and effectively prepare for classes, the course instructor will maintain a class
schedule with reading assignments and presentation slides.

% During the class sessions students will also be required to
% download, use, and modify programs, diagrams, and data sets that are made available through the course Web site.
% Students who are not comfortable with compiling, editing, and running Java programs should see the course instructor.

\subsubsection*{Seeking Assistance}

Students who are struggling to understand the knowledge and skills developed in a class or laboratory session are
encourage to seek assistance from the course instructor. Throughout the semester, students should, within the bounds of
the Honor Code, ask and answer questions on the Slack site for our course; please request assistance from the instructor
first through Slack before sending an email. Students who need the course instructor's assistance must schedule a
meeting through his Web site and come to the meeting with all of the details needed to discuss their question.

\subsubsection*{Using Email}

Although we will primarily use Slack for class communication, I will sometimes use email to send announcements about
important matters such as changes in the schedule. It is your responsibility to check your email at least once a day and to
ensure that you can reliably send and receive emails. This class policy is based on the statement about the use of email that
appears in {\em The Compass}, the College's student handbook; please see the instructor if you do not have this
handbook.

% Using your Allegheny College email address, I will sometimes send out class announcements about matters such as
% assignment clarifications or changes in the schedule. It is your responsibility to check your email at least once a day
% and to ensure that you can reliably send and receive emails. This class policy is based on the following statement in
% {\em The Compass}, the college's student handbook.

% \vspace*{-.1in}
% \begin{quote}
% ``The use of email is a primary method of communication on campus. \ldots
% All students are provided with a campus email account and address while
% enrolled at Allegheny and are expected to check the account on a regular
% basis.''
% \end{quote}
% \vspace*{-.15in}

\subsubsection*{Disability Services}

The Americans with Disabilities Act (ADA) is a federal anti-discrimination statute that provides comprehensive civil
rights protection for persons with disabilities.  Among other things, this legislation requires all students with
disabilities be guaranteed a learning environment that provides for reasonable accommodation of their disabilities.
Students with disabilities who believe they may need accommodations in this class are encouraged to contact Disability
Services at 332-2898.  Disability Services is part of the Learning Commons and is located in Pelletier Library.
Please do this as soon as possible to ensure that approved accommodations are implemented in a timely fashion.

\subsubsection*{Honor Code}

The Academic Honor Program that governs the entire academic program at Allegheny College is described in the Allegheny
Academic Bulletin.  The Honor Program applies to all work that is submitted for academic credit or to meet non-credit
requirements for graduation at Allegheny College.  This includes all work assigned for this class (e.g., examinations,
laboratory assignments, and the final project).  All students enrolled in the College will work under the Honor Program.

% Each student who has matriculated at the College has acknowledged the following pledge:

% \vspace*{-.1in}
% \begin{quote}
% I hereby recognize and pledge to fulfill my responsibilities, as defined in the Honor Code, and to maintain the
% integrity of both myself and the College community as a whole.
% \end{quote}
% \vspace*{-.15in}

\noindent It is understood that an important part of the learning process in any course, and particularly one in
computer science, derives from thoughtful discussions with teachers and fellow students.  While it is acceptable for
students in this class to discuss their work with their classmates, deliverables that are nearly identical to the work
of others will be taken as evidence of violating the \mbox{Honor Code}.

\subsection*{Welcome to a Software Engineering Adventure}

In reference to software, Frederick Brooks, Jr.\ wrote in Chapter One of MMM, ``The magic of myth and legend has come
true in our time.'' Software is a pervasive aspect of our society that changes how we think and act.  High quality
software also has the potential to positively influence the lives of people. Moreover, the specification, design,
implementation, testing, maintenance, and documentation of software are exciting and rewarding activities!  At the start
of this class, I invite you to pursue, with great enthusiasm and vigor, this adventure in software engineering.

\end{document}
